---
title: "CMVTRAP analysis"
output:
  html_document:
    keep_md: yes
    theme: united
    toc: yes
  pdf_document:
    toc: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(edgeR)
library(SummarizedExperiment)
library(SEtools)
library(plgINS)
library(ggplot2)
library(sva)
library(DESeq2)
library(topGO)
source("BioStatsfunctions_LvZ.R")
```

# Analysis of CMV TRAP
## Analysis of CMV TRAP data with full model (unbound vs bound)

```{r}
SE <- readRDS("../data/CMVTRAP.vHC.SE.rds")
SE$SampleProcessing <- as.factor(SE$SampleProcessing)
SE$SampleProcessing <- relevel(SE$SampleProcessing, "WholeRNA")
SE <-SE[,order(SE$SampleProcessing,SE$Condition)]

design <- model.matrix(~SE$SampleProcessing * SE$Condition)

y <- DGEList(counts=assays(SE)$counts)
y <- calcNormFactors(y)
y <- estimateDisp(y,design)
#filter out genes that are low expressed
y <- y[filterByExpr(y, design),]

Results <- list()
fit <- glmQLFit(y,design)
for(i in colnames(design)[-1]){
  Results[[i]] <- glmQLFTest(fit, i)
}

#Plot these genes in the
for(i in names(Results)){
  sg <- rownames(topTags(Results[[i]],n = 1000, p.value = 0.05))
  if(length(sg) > 1){
print(sehm(SE,genes = sg, 
     gaps_at = c("Experiment","SampleProcessing","Region","Sex"),
     do.scale = T, 
     anno_columns = c("Experiment","Sex","Condition","TimePoint", "SampleProcessing","Region"), 
     cluster_rows = T, main = paste("sig. in",i ,sep ="_"),
     assayName = "logcpm"))
print(topTags(Results[[i]],p.value = 0.05, n = 30))
  }else{
  print(paste(i, "had no significant entries", sep = " "))
  print(topTags(Results[[i]], n = 30))
}
}

VolcanoPlotsEdgeR(Results)

gene_universe <- p.adjust(Results$`SE$SampleProcessingTRAP`$table$PValue,method = "BH" ) <= 0.05
gene_universe <- as.numeric(gene_universe)
gene_universe <- factor(gene_universe)
names(gene_universe) <- rownames(Results$`SE$SampleProcessingTRAP`$table)

go_table <- list()
for(i in c("CC","BP","MF")){
go_data <- new("topGOdata",
               ontology = i,
               allGenes = gene_universe,
               nodeSize = 10,
               annotationFun = annFUN.org,
               mapping = "org.Mm.eg",
               ID = "symbol")

go_test <- runTest(go_data, algorithm = "weight01", statistic = "fisher")

allGO = usedGO(object = go_data)
go_table[[i]] <- GenTable(go_data, weightFisher = go_test,
                     orderBy = "weightFisher", ranksOf = "weightFisher",
                     topNodes = length(allGO))
print(paste("topGO ontology type:", i, sep = " "))
print(head(go_table[[i]],20))
}

```

## Independent analysis in bound fraction

Let's start by plotting stress genes from the timeseries in the new data set and correct for experiment and sample processing. we set CMV TRAP as the baseline

```{r}
SE <- readRDS("../data/CMVTRAP.vHC.SE.rds")
SE$SampleProcessing <- as.factor(SE$SampleProcessing)
SE$SampleProcessing <- relevel(SE$SampleProcessing, "WholeRNA")
SE <-SE[,order(SE$SampleProcessing,SE$Condition)]
SE <- SE[,SE$SampleProcessing == "TRAP"]

y <- DGEList(counts=assays(SE)$counts, group = SE$Condition)
y <- calcNormFactors(y)
y <- estimateDisp(y)
#filter out genes that are low expressed
y <- y[filterByExpr(y),]

Results <- list()
Results[["Swim"]] <- exactTest(y)


#Plot these genes in the
for(i in names(Results)){
  sg <-  rownames(topTags(Results[[i]],p.value = 0.05, n = 1000))
  if(length(sg) > 1){
print(sehm(SE,genes = sg, 
     gaps_at = c("Experiment","SampleProcessing","Region","Sex"),
     do.scale = T, 
     anno_columns = c("Experiment","Sex","Condition","TimePoint", "SampleProcessing","Region"), 
     cluster_rows = T, main = paste("sig. in",i ,sep ="_"),
     assayName = "logcpm"))
print(topTags(Results[[i]],p.value = 0.05, n = 30))
  }else{
  print(paste(i, "had no significant entries", sep = " "))
  print(topTags(Results[[i]], n = 30))
}
}

VolcanoPlotsEdgeR(Results)

```

## Independent analysis in unbound fraction

Let's start by plotting stress genes from the timeseries in the new data set and correct for experiment and sample processing. we set CMV TRAP as the baseline

```{r}
SE <- readRDS("../data/CMVTRAP.vHC.SE.rds")
SE$SampleProcessing <- as.factor(SE$SampleProcessing)
SE$SampleProcessing <- relevel(SE$SampleProcessing, "WholeRNA")
SE <-SE[,order(SE$SampleProcessing,SE$Condition)]
SE <- SE[,SE$SampleProcessing == "WholeRNA"]

y <- DGEList(counts=assays(SE)$counts, group = SE$Condition)
y <- calcNormFactors(y)
y <- estimateDisp(y)
#filter out genes that are low expressed
y <- y[filterByExpr(y),]

Results <- list()
Results[["Swim"]] <- exactTest(y)


#Plot these genes in the
for(i in names(Results)){
  sg <-  rownames(topTags(Results[[i]],p.value = 0.05, n = 1000))
  if(length(sg) > 1){
print(sehm(SE,genes = sg, 
     gaps_at = c("Experiment","SampleProcessing","Region","Sex"),
     do.scale = T, 
     anno_columns = c("Experiment","Sex","Condition","TimePoint", "SampleProcessing","Region"), 
     cluster_rows = T, main = paste("sig. in",i ,sep ="_"),
     assayName = "logcpm"))
print(topTags(Results[[i]],p.value = 0.05, n = 30))
  }else{
  print(paste(i, "had no significant entries", sep = " "))
  print(topTags(Results[[i]], n = 30))
}
}

VolcanoPlotsEdgeR(Results)
```

## Analysis of CMV TRAP data using interaction term after removing processing effects

Let's start by plotting stress genes from the timeseries in the new data set and correct for experiment and sample processing. we set CMV TRAP as the baseline

```{r}
SE <- readRDS("../data/CMVTRAP.vHC.SE.rds")
SE$SampleProcessing <- as.factor(SE$SampleProcessing)
SE$SampleProcessing <- relevel(SE$SampleProcessing, "WholeRNA")
SE <-SE[,order(SE$SampleProcessing,SE$Condition)]

SE <- dosvacor(SE,form = ~SampleProcessing * Condition, form0 = ~SampleProcessing)
design <- model.matrix(~SE$SV1 + SE$SV2 + SE$SV3 + SE$Condition * SE$SampleProcessing)

y <- DGEList(counts=assays(SE)$counts)
y <- calcNormFactors(y)
y <- estimateDisp(y,design)
#filter out genes that are low expressed
y <- y[filterByExpr(y, design),]

Results <- list()
fit <- glmQLFit(y,design)
for(i in colnames(design)[c(5,7)]){
  Results[[i]] <- glmQLFTest(fit, i)
}
Results[["FullModel"]] <-  glmQLFTest(fit, colnames(design)[c(5,7)])

#Define significant genes
select <- rownames(topTags(Results[["FullModel"]],p.value = 0.05, n = 1000))


#Plot these genes in the
for(i in names(Results)){
  sg <- select[Results[[i]]$table[select,]$PValue <= 0.05]
  if(length(sg) > 1){
print(sehm(SE,genes = sg, 
     gaps_at = c("Experiment","SampleProcessing","Region","Sex"),
     do.scale = T, 
     anno_columns = c("Experiment","Sex","Condition","TimePoint", "SampleProcessing","Region"), 
     cluster_rows = T, main = paste("sig. in",i ,sep ="_"),
     assayName = "corrected"))
print(topTags(Results[[i]],p.value = 0.05, n = 30))
  }else{
  print(paste(i, "had no significant entries", sep = " "))
  print(topTags(Results[[i]], n = 30))
}
}

VolcanoPlotsEdgeR(Results[-3])
```

# Meta analysis of all TRAP data

## Analysis of combined data (additive)

Let's start by plotting stress genes from the timeseries in the new data set and correct for experiment and sample processing

```{r}
SE <- readRDS("../data/CMVTRAP.vHC.SE.rds")
SE$SampleNumber <- SE$SampleNr
SE$SampleNr <- NULL
SE$Block <- "CMVTRAP_Block1"
TS1 <- readRDS("../data/TimeSeriesFST.vHC.SE.rds")
TS1$CageNumber <- NULL
IT <- readRDS("../data/VIAAT-TRAP.vHC.SE.rds")
IT$SampleNumber <- NA
ET1 <- readRDS("../data/CAMK2A-TRAP.vHC.SE.rds")
ET1$TimePoint <- ET1$Timepoint
ET1$Timepoint <- NULL
ET2 <- readRDS("../data/CAMK2A-TRAP_Exp2.vHC.SE.rds")
ET2$TimePoint <- ET2$Timepoint
ET2$Timepoint <- NULL

genes <- intersect(names(SE),names(TS1))
genes <- intersect(genes,names(IT))
genes <- intersect(genes,names(ET1))
genes <- intersect(genes,names(ET2))

ts_sg <- read.table("../metadata/TimeSeries_StatisticalResults.csv", sep = ";", header = T)
ts_sg <- unique(ts_sg[(ts_sg$adj.P.Val <= 0.05 & ts_sg$TimePoint == "Swim_45min") ,"ShortName"])
CB <- cbind(TS1[genes,],
            SE[genes,],
            IT[genes,],
            ET1[genes,],
            ET2[genes,])
CB$SampleProcessing2 <- paste(CB$Experiment,CB$SampleProcessing, sep = "_")
CB <- CB[,order(CB$Experiment, CB$SampleProcessing, CB$Condition, CB$TimePoint)]
CB <- dosvacor(CB, form = ~SampleProcessing2 + TimePoint, form0 = ~SampleProcessing2)
sehm(CB,genes = ts_sg, 
     gaps_at = c("Experiment","SampleProcessing","Region","Sex"),
     do.scale = T, 
     anno_columns = c("Experiment","Sex","Condition","TimePoint", "SampleProcessing","Region"), 
     cluster_rows = T, main = "sig. genes from TS, batch corrected with all TRAP",
     assayName = "corrected")

CB2 <- CB[,CB$TimePoint %in% c("0min","45min")]

design <- model.matrix(~CB2$SV1 + CB2$SV2 + CB2$SV3 + CB2$SV4 + CB2$SV5 + CB2$SV6 + CB2$SV7 + CB2$SV8 + CB2$SV9 + CB2$SV10 + CB2$SV11 + CB2$SV12 + CB2$SV13 + CB2$SV14 + CB2$SV15 + CB2$SV16 + CB2$SV17 + CB2$SV18 + CB2$SV19 + CB2$SV20 + CB2$SV21 + CB2$SV22 + CB2$SV23 + CB2$Condition)

y <- DGEList(counts=assays(CB2)$counts)
y <- calcNormFactors(y)
y <- estimateDisp(y,design)
#filter out genes that are low expressed
y <- y[filterByExpr(y, design),]

Results <- list()
fit <- glmQLFit(y,design)
for(i in colnames(design)[-c(1:24)]){
  Results[[i]] <- glmQLFTest(fit, i)
}

for(i in names(Results)){
  sg <- rownames(topTags(Results[[i]],p.value = 0.05, n = 1000))
  if(length(sg) > 1){
print(sehm(CB2,genes = sg, 
     gaps_at = c("Experiment","SampleProcessing","Region","Sex"),
     do.scale = T, 
     anno_columns = c("Experiment","Sex","Condition","TimePoint", "SampleProcessing","Region"), 
     cluster_rows = T, main = "sig. genes (swim 45min) from combined analysis",
     assayName = "corrected"))
print(topTags(Results[[i]],p.value = 0.05, n = 30))
  }else{
  print(paste(i, "had no significant entries", sep = " "))
  print(topTags(Results[[i]], n = 30))
}
}
```
## Analysis of combined data (additive) with randomization as control
There seems to be a huge agreement between the data for some 45 minute genes. As a sanity check, lets randomly re-assign the treatment in one data-set (inhibitory TRAP) and make sure that this effect disappears in this data-set

```{r}
SE <- readRDS("../data/CMVTRAP.vHC.SE.rds")
SE$SampleNumber <- SE$SampleNr
SE$SampleNr <- NULL
SE$Block <- "CMVTRAP_Block1"
TS1 <- readRDS("../data/TimeSeriesFST.vHC.SE.rds")
TS1$CageNumber <- NULL
IT <- readRDS("../data/VIAAT-TRAP.vHC.SE.rds")
IT$SampleNumber <- NA
IT$TimePoint <- sample(IT$TimePoint)
ET1 <- readRDS("../data/CAMK2A-TRAP.vHC.SE.rds")
ET1$TimePoint <- ET1$Timepoint
ET1$Timepoint <- NULL
ET2 <- readRDS("../data/CAMK2A-TRAP_Exp2.vHC.SE.rds")
ET2$TimePoint <- ET2$Timepoint
ET2$Timepoint <- NULL

genes <- intersect(names(SE),names(TS1))
genes <- intersect(genes,names(IT))
genes <- intersect(genes,names(ET1))
genes <- intersect(genes,names(ET2))

ts_sg <- read.table("../metadata/TimeSeries_StatisticalResults.csv", sep = ";", header = T)
ts_sg <- unique(ts_sg[(ts_sg$adj.P.Val <= 0.05 & ts_sg$TimePoint == "Swim_45min") ,"ShortName"])
CB <- cbind(TS1[genes,],
            SE[genes,],
            IT[genes,],
            ET1[genes,],
            ET2[genes,])
CB$SampleProcessing2 <- paste(CB$Experiment,CB$SampleProcessing, sep = "_")
CB <- CB[,order(CB$Experiment, CB$SampleProcessing, CB$TimePoint)]
CB <- dosvacor(CB, form = ~SampleProcessing2 + TimePoint, form0 = ~SampleProcessing2)
sehm(CB,genes = ts_sg, 
     gaps_at = c("Experiment","SampleProcessing","Region","Sex"),
     do.scale = T, 
     anno_columns = c("Experiment","Sex","Condition","TimePoint", "SampleProcessing","Region"), 
     cluster_rows = T, main = "sig. genes from TS, batch corrected, 1 experiment randomized",
     assayName = "corrected")

CB2 <- CB[,CB$TimePoint %in% c("0min","45min")]

design <- model.matrix(~CB2$SV1 + CB2$SV2 + CB2$SV3 + CB2$SV4 + CB2$SV5 + CB2$SV6 + CB2$SV7 + CB2$SV8 + CB2$SV9 + CB2$SV10 + CB2$SV11 + CB2$SV12 + CB2$SV13 + CB2$SV14 + CB2$SV15 + CB2$SV16 + CB2$SV17 + CB2$SV18 + CB2$SV19 + CB2$SV20 + CB2$SV21 + CB2$SV22 + CB2$SV23 + CB2$Condition)

y <- DGEList(counts=assays(CB2)$counts)
y <- calcNormFactors(y)
y <- estimateDisp(y,design)
#filter out genes that are low expressed
y <- y[filterByExpr(y, design),]

Results <- list()
fit <- glmQLFit(y,design)
for(i in colnames(design)[-c(1:24)]){
  Results[[i]] <- glmQLFTest(fit, i)
}

for(i in names(Results)){
  sg <- rownames(topTags(Results[[i]],p.value = 0.05, n = 1000))
  if(length(sg) > 1){
print(sehm(CB2,genes = sg, 
     gaps_at = c("Experiment","SampleProcessing","Region","Sex"),
     do.scale = T, 
     anno_columns = c("Experiment","Sex","Condition","TimePoint", "SampleProcessing","Region"), 
     cluster_rows = T, main = "sig. genes (swim 45min) from combined analysis",
     assayName = "corrected"))
print(topTags(Results[[i]],p.value = 0.05, n = 30))
  }else{
  print(paste(i, "had no significant entries", sep = " "))
  print(topTags(Results[[i]], n = 30))
}
}
```

As we can see, the effect is not introduced by the SVA correction, but really there!
The next working hypothesis is that SVA can correct for these effects as long as they are present in experiments / sample processing. Lets test this by pairing different experiments with the TimeSeries and applying SVA

### SVA pairing of inhibitory TRAP + TS

```{r}
TS1 <- readRDS("../data/TimeSeriesFST.vHC.SE.rds")
TS1$CageNumber <- NULL
IT <- readRDS("../data/VIAAT-TRAP.vHC.SE.rds")
IT$SampleNumber <- NA


genes <- intersect(names(TS1),names(IT))

ts_sg <- read.table("../metadata/TimeSeries_StatisticalResults.csv", sep = ";", header = T)
ts_sg <- unique(ts_sg[(ts_sg$adj.P.Val <= 0.05 & ts_sg$TimePoint == "Swim_45min") ,"ShortName"])
CB <- cbind(TS1[genes,],
            IT[genes,])
CB$SampleProcessing2 <- paste(CB$Experiment,CB$SampleProcessing, sep = "_")
CB <- CB[,order(CB$Experiment, CB$SampleProcessing, CB$TimePoint)]
CB <- dosvacor(CB, form = ~SampleProcessing2 + TimePoint, form0 = ~SampleProcessing2)
sehm(CB,genes = ts_sg, 
     gaps_at = c("Experiment","SampleProcessing","Region","Sex"),
     do.scale = T, 
     anno_columns = c("Experiment","Sex","Condition","TimePoint", "SampleProcessing","Region"), 
     cluster_rows = T, main = "sig genes from TS, in TS and inhibitory TRAP",
     assayName = "corrected")

CB2 <- CB[,CB$TimePoint %in% c("0min","45min")]

design <- model.matrix(~CB2$SV1 + CB2$SV2 + CB2$SV3 + CB2$SV4 + CB2$SV5 + CB2$SV6 + CB2$SV7 + CB2$SV8 + CB2$SV9 + CB2$SV10 + CB2$SV11 + CB2$SV12 + CB2$Condition)

y <- DGEList(counts=assays(CB2)$counts)
y <- calcNormFactors(y)
y <- estimateDisp(y,design)
#filter out genes that are low expressed
y <- y[filterByExpr(y, design),]

Results <- list()
fit <- glmQLFit(y,design)
for(i in colnames(design)[-c(1:13)]){
  Results[[i]] <- glmQLFTest(fit, i)
}

for(i in names(Results)){
  sg <- rownames(topTags(Results[[i]],p.value = 0.05, n = 1000))
  if(length(sg) > 1){
print(sehm(CB2,genes = sg, 
     gaps_at = c("Experiment","SampleProcessing","Region","Sex"),
     do.scale = T, 
     anno_columns = c("Experiment","Sex","Condition","TimePoint", "SampleProcessing","Region"), 
     cluster_rows = T, main = "sig. genes (swim 45min) in TS + inhibitory TRAP",
     assayName = "corrected"))
print(topTags(Results[[i]],p.value = 0.05, n = 30))
  }else{
  print(paste(i, "had no significant entries", sep = " "))
  print(topTags(Results[[i]], n = 30))
}
}
```
### SVA pairing of excitatory TRAP + TS

```{r}
TS1 <- readRDS("../data/TimeSeriesFST.vHC.SE.rds")
TS1$CageNumber <- NULL
ET1 <- readRDS("../data/CAMK2A-TRAP.vHC.SE.rds")
ET1$TimePoint <- ET1$Timepoint
ET1$Timepoint <- NULL
ET2 <- readRDS("../data/CAMK2A-TRAP_Exp2.vHC.SE.rds")
ET2$TimePoint <- ET2$Timepoint
ET2$Timepoint <- NULL

genes <- intersect(names(TS1),names(ET1))
genes <- intersect(genes,names(ET2))

ts_sg <- read.table("../metadata/TimeSeries_StatisticalResults.csv", sep = ";", header = T)
ts_sg <- unique(ts_sg[(ts_sg$adj.P.Val <= 0.05 & ts_sg$TimePoint == "Swim_45min") ,"ShortName"])
CB <- cbind(TS1[genes,],
            ET1[genes,],
            ET2[genes,])
CB$SampleProcessing2 <- paste(CB$Experiment,CB$SampleProcessing, sep = "_")
CB <- CB[,order(CB$Experiment, CB$SampleProcessing, CB$TimePoint)]
CB <- dosvacor(CB, form = ~SampleProcessing2 + TimePoint, form0 = ~SampleProcessing2)
sehm(CB,genes = ts_sg, 
     gaps_at = c("Experiment","SampleProcessing","Region","Sex"),
     do.scale = T, 
     anno_columns = c("Experiment","Sex","Condition","TimePoint", "SampleProcessing","Region"), 
     cluster_rows = T, main = "sig genes from TS, in TS and excitatory TRAP",
     assayName = "corrected")

#Let run a differential analysis on the 45min and 1h30min timepoints
CB2 <- CB[,CB$TimePoint %in% c("0min","45min","1h30min")]
CB2$TimePoint <- factor(CB2$TimePoint,levels = c("0min","45min","1h30min"))

design <- model.matrix(~CB2$SV1 + CB2$SV2 + CB2$SV3 + CB2$SV4 + CB2$SV5 + CB2$SV6 + CB2$SV7 + CB2$SV8 + CB2$SV9 + CB2$SV10 + CB2$SV11 + CB2$SV12 + CB2$SV13 + CB2$SV14 + CB2$SV15 + CB2$SV16 + CB2$SV17 + CB2$SV18 + CB2$SV19 + CB2$SV20 + CB2$SV21 + CB2$SV22 + CB2$SV23 + CB2$SV24 + CB2$SV25 + CB2$TimePoint)

y <- DGEList(counts=assays(CB2)$counts)
y <- calcNormFactors(y)
y <- estimateDisp(y,design)
#filter out genes that are low expressed
y <- y[filterByExpr(y, design),]

Results <- list()
fit <- glmQLFit(y,design)
Results[["ConditionSwim"]] <- glmQLFTest(fit, colnames(design)[27:28])

for(i in names(Results)){
  sg <- rownames(topTags(Results[[i]],p.value = 0.05, n = 1000))
  if(length(sg) > 1){
print(sehm(CB2,genes = sg, 
     gaps_at = c("Experiment","SampleProcessing","Region","Sex"),
     do.scale = T, 
     anno_columns = c("Experiment","Sex","Condition","TimePoint", "SampleProcessing","Region"), 
     cluster_rows = T, main = "sig. genes (swim 45min) in TS and excitatory TRAP",
     assayName = "corrected"))
print(topTags(Results[[i]],p.value = 0.05, n = 30))
  }else{
  print(paste(i, "had no significant entries", sep = " "))
  print(topTags(Results[[i]], n = 30))
}
}
```

### SVA pairing of CMV TRAP (bound and unbound) + TS

```{r}
TS1 <- readRDS("../data/TimeSeriesFST.vHC.SE.rds")
TS1$CageNumber <- NULL
SE <- readRDS("../data/CMVTRAP.vHC.SE.rds")
SE$SampleNumber <- SE$SampleNr
SE$SampleNr <- NULL
SE$Block <- "CMVTRAP_Block1"

genes <- intersect(names(TS1),names(SE))

ts_sg <- read.table("../metadata/TimeSeries_StatisticalResults.csv", sep = ";", header = T)
ts_sg <- unique(ts_sg[(ts_sg$adj.P.Val <= 0.05 & ts_sg$TimePoint == "Swim_45min") ,"ShortName"])
CB <- cbind(TS1[genes,],
            SE[genes,])
CB$SampleProcessing2 <- paste(CB$Experiment,CB$SampleProcessing, sep = "_")
CB <- CB[,order(CB$Experiment, CB$SampleProcessing, CB$TimePoint)]
CB <- dosvacor(CB, form = ~SampleProcessing2 + TimePoint, form0 = ~SampleProcessing2)
sehm(CB,genes = ts_sg, 
     gaps_at = c("Experiment","SampleProcessing","Region","Sex"),
     do.scale = T, 
     anno_columns = c("Experiment","Sex","Condition","TimePoint", "SampleProcessing","Region"), 
     cluster_rows = T, main = "sig genes from TS, in TS and CMV TRAP",
     assayName = "corrected")

CB2 <- CB[,CB$TimePoint %in% c("0min","45min")]

design <- model.matrix(~CB2$SV1 + CB2$SV2 + CB2$SV3 + CB2$SV4 + CB2$SV5 + CB2$SV6 + CB2$SV7 + CB2$SV8 + CB2$SV9 + CB2$SV10 + CB2$SV11 + CB2$SV12 + CB2$SV13 + CB2$SV14 + CB2$SV15 + CB2$SV16 + CB2$SV17 + CB2$Condition)

y <- DGEList(counts=assays(CB2)$counts)
y <- calcNormFactors(y)
y <- estimateDisp(y,design)
#filter out genes that are low expressed
y <- y[filterByExpr(y, design),]

Results <- list()
fit <- glmQLFit(y,design)
for(i in colnames(design)[-c(1:18)]){
  Results[[i]] <- glmQLFTest(fit, i)
}

for(i in names(Results)){
  sg <- rownames(topTags(Results[[i]],p.value = 0.05, n = 1000))
  if(length(sg) > 1){
print(sehm(CB2,genes = sg, 
     gaps_at = c("Experiment","SampleProcessing","Region","Sex"),
     do.scale = T, 
     anno_columns = c("Experiment","Sex","Condition","TimePoint", "SampleProcessing","Region"), 
     cluster_rows = T, main = "sig genes (swim 45min) in TS and CMV TRAP (bound and unbound)",
     assayName = "corrected"))
print(topTags(Results[[i]],p.value = 0.05, n = 30))
  }else{
  print(paste(i, "had no significant entries", sep = " "))
  print(topTags(Results[[i]], n = 30))
}
}
```

## Analysis of combined data (with interaction) at 45min

Let's start by plotting stress genes from the timeseries in the new data set and correct for experiment and sample processing

```{r}
SE <- readRDS("../data/CMVTRAP.vHC.SE.rds")
SE$SampleNumber <- SE$SampleNr
SE$SampleNr <- NULL
SE$Block <- "CMVTRAP_Block1"
TS1 <- readRDS("../data/TimeSeriesFST.vHC.SE.rds")
TS1$CageNumber <- NULL
TS1$Experiment <- "TimeSeries"
IT <- readRDS("../data/VIAAT-TRAP.vHC.SE.rds")
IT$SampleNumber <- NA
IT$Experiment <- "VIAAT"
ET1 <- readRDS("../data/CAMK2A-TRAP.vHC.SE.rds")
ET1$TimePoint <- ET1$Timepoint
ET1$Timepoint <- NULL
ET1$Experiment <- "CAMK2A"
ET2 <- readRDS("../data/CAMK2A-TRAP_Exp2.vHC.SE.rds")
ET2$TimePoint <- ET2$Timepoint
ET2$Timepoint <- NULL
ET2$Experiment <- "CAMK2A"

genes <- intersect(names(SE),names(TS1))
genes <- intersect(genes,names(IT))
genes <- intersect(genes,names(ET1))
genes <- intersect(genes,names(ET2))

ts_sg <- read.table("../metadata/TimeSeries_StatisticalResults.csv", sep = ";", header = T)
ts_sg <- unique(ts_sg[(ts_sg$adj.P.Val <= 0.05 & ts_sg$TimePoint == "Swim_45min") ,"ShortName"])
CB <- cbind(TS1[genes,],
            SE[genes,],
            IT[genes,],
            ET1[genes,],
            ET2[genes,])
CB$SampleProcessing2 <- paste(CB$Experiment,CB$SampleProcessing, sep = "_")
CB$Experiment <- as.factor(CB$Experiment)
CB$Experiment <- relevel(CB$Experiment,"TimeSeries")
CB <- CB[,order(CB$Experiment, CB$SampleProcessing, CB$Condition, CB$TimePoint)]
CB <- CB[,CB$TimePoint %in% c("0min","45min")]
CB$TimePoint <- factor(CB$TimePoint,levels =  c("0min","45min"))
CB$SampleProcessing2 <- as.factor(CB$SampleProcessing2)
CB$SampleProcessing2 <- relevel(CB$SampleProcessing2,"TimeSeries_WholeRNA")

CB <- dosvacor(CB, form = ~SampleProcessing2 * TimePoint, form0 = ~SampleProcessing2)
sehm(CB,genes = ts_sg, 
     gaps_at = c("Experiment","SampleProcessing","Region","Sex"),
     do.scale = T, 
     anno_columns = c("Experiment","Sex","Condition","TimePoint", "SampleProcessing","Region"), 
     cluster_rows = T, main = "sig. genes from TS, batch corrected in TS and all TRAP",
     assayName = "corrected")

CB2 <- CB[,CB$TimePoint %in% c("0min","45min")]

design <- model.matrix(~CB2$SV1 + CB2$SV2 + CB2$SV3 + CB2$SV4 + CB2$SV5 + CB2$SV6 + CB2$SV7 + CB2$SV8 + CB2$SV9 + CB2$SV10 + CB2$SV11 + CB2$SV12 +CB$Condition * CB$SampleProcessing2)

y <- DGEList(counts=assays(CB2)$counts)
y <- calcNormFactors(y)
y <- estimateDisp(y,design)
#filter out genes that are low expressed
y <- y[filterByExpr(y, design),]

Results <- list()
fit <- glmQLFit(y,design)
for(i in colnames(design)[c(14,19:22)]){
  Results[[i]] <- glmQLFTest(fit, i)
}
Results[["FullModel"]] <-  glmQLFTest(fit, colnames(design)[c(14,19:22)])

#Define significant genes
select <- rownames(topTags(Results[["FullModel"]],p.value = 0.05, n = 1000))

#Plot these genes in the
for(i in names(Results)){
  sg <- select[Results[[i]]$table[select,]$PValue <= 0.05]
  if(length(sg) > 1){
print(sehm(CB2,genes = sg, 
     gaps_at = c("Experiment","SampleProcessing","Region","Sex"),
     do.scale = T, 
     anno_columns = c("Experiment","Sex","Condition","TimePoint", "SampleProcessing","Region"), 
     cluster_rows = T, main = paste("sig genes in",i ,sep ="_"),
     assayName = "corrected"))
print(topTags(Results[[i]],p.value = 0.05, n = 30))
  }else{
  print(paste(i, "had no significant entries", sep = " "))
  print(topTags(Results[[i]], n = 30))
}
}

tags <- topTags(Results[["FullModel"]],p.value = 0.05, n = 20)$table

radial.df <- data.frame(name = rownames(tags), logFC = tags$logFC.CB.ConditionSwim, Experiment = "TimeSeries")
radial.df <- rbind(radial.df, data.frame(name = rownames(tags), logFC = tags$logFC.CB.ConditionSwim + tags$logFC.CB.ConditionSwim.CB.SampleProcessing2CAMK2A_TRAP, Experiment = "Camk2a TRAP"))
radial.df <- rbind(radial.df, data.frame(name = rownames(tags), logFC = tags$logFC.CB.ConditionSwim + tags$logFC.CB.ConditionSwim.CB.SampleProcessing2CMV.TRAP_TRAP, Experiment = "CMV TRAP"))
radial.df <- rbind(radial.df, data.frame(name = rownames(tags), logFC = tags$logFC.CB.ConditionSwim + tags$logFC.CB.ConditionSwim.CB.SampleProcessing2VIAAT_TRAP, Experiment = "Viaat TRAP"))
radial.df$direction = ifelse(radial.df$logFC >= 0, "up-regulated", "down-regulated")
radial.df$logFC <- abs(radial.df$logFC)
radial.df$Experiment <- as.factor(radial.df$Experiment)
radial.df$Experiment <- relevel(radial.df$Experiment, "TimeSeries")

ggplot(radial.df, aes(x=name, y=logFC, fill=name, color = direction)) +
    geom_bar(stat='identity') + facet_wrap(~Experiment) + coord_polar() + scale_color_manual(values= c("red","black")) + theme_bw() + ggtitle("top 20 genes")

```

## Analysis of combined data (with interaction) at 45min without timeseries

Let's start by plotting stress genes from the timeseries in the new data set and correct for experiment and sample processing. we set CMV TRAP as the baseline

```{r}
SE <- readRDS("../data/CMVTRAP.vHC.SE.rds")
SE$SampleNumber <- SE$SampleNr
SE$SampleNr <- NULL
SE$Block <- "CMVTRAP_Block1"
IT <- readRDS("../data/VIAAT-TRAP.vHC.SE.rds")
IT$SampleNumber <- NA
IT$Experiment <- "VIAAT"
ET1 <- readRDS("../data/CAMK2A-TRAP.vHC.SE.rds")
ET1$TimePoint <- ET1$Timepoint
ET1$Timepoint <- NULL
ET1$Experiment <- "CAMK2A"
ET2 <- readRDS("../data/CAMK2A-TRAP_Exp2.vHC.SE.rds")
ET2$TimePoint <- ET2$Timepoint
ET2$Timepoint <- NULL
ET2$Experiment <- "CAMK2A"

genes <- intersect(names(SE),names(IT))
genes <- intersect(genes,names(ET1))
genes <- intersect(genes,names(ET2))

ts_sg <- read.table("../metadata/TimeSeries_StatisticalResults.csv", sep = ";", header = T)
ts_sg <- unique(ts_sg[(ts_sg$adj.P.Val <= 0.05 & ts_sg$TimePoint == "Swim_45min") ,"ShortName"])
CB <- cbind(SE[genes,],
            IT[genes,],
            ET1[genes,],
            ET2[genes,])
CB$SampleProcessing2 <- paste(CB$Experiment,CB$SampleProcessing, sep = "_")
CB$Experiment <- as.factor(CB$Experiment)
CB$Experiment <- relevel(CB$Experiment,"CMV-TRAP")
CB <- CB[,order(CB$Experiment, CB$SampleProcessing, CB$Condition, CB$TimePoint)]
CB <- CB[,CB$TimePoint %in% c("0min","45min")]
CB$TimePoint <- factor(CB$TimePoint,levels =  c("0min","45min"))
CB$SampleProcessing2 <- as.factor(CB$SampleProcessing2)
CB$SampleProcessing2 <- relevel(CB$SampleProcessing2,"CMV-TRAP_TRAP")

CB <- dosvacor(CB, form = ~SampleProcessing2 * TimePoint, form0 = ~SampleProcessing2)
sehm(CB,genes = ts_sg, 
     gaps_at = c("Experiment","SampleProcessing","Region","Sex"),
     do.scale = T, 
     anno_columns = c("Experiment","Sex","Condition","TimePoint", "SampleProcessing","Region"), 
     cluster_rows = T, main = "sig. genes from TS, batch corrected in all TRAP",
     assayName = "corrected")

CB2 <- CB[,CB$TimePoint %in% c("0min","45min")]

design <- model.matrix(~CB2$SV1 + CB2$SV2 + CB2$SV3 + CB2$SV4 + CB2$SV5 + CB2$SV6 + CB2$SV7 + CB2$SV8 +CB$Condition * CB$SampleProcessing2)

y <- DGEList(counts=assays(CB2)$counts)
y <- calcNormFactors(y)
y <- estimateDisp(y,design)
#filter out genes that are low expressed
y <- y[filterByExpr(y, design),]

Results <- list()
fit <- glmQLFit(y,design)
for(i in colnames(design)[c(10,14:16)]){
  Results[[i]] <- glmQLFTest(fit, i)
}
Results[["FullModel"]] <-  glmQLFTest(fit, colnames(design)[c(10,14:16)])

#Define significant genes
select <- rownames(topTags(Results[["FullModel"]],p.value = 0.05, n = 1000))

#Plot these genes in the
for(i in names(Results)){
  sg <- select[Results[[i]]$table[select,]$PValue <= 0.05]
  if(length(sg) > 1){
print(sehm(CB2,genes = sg, 
     gaps_at = c("Experiment","SampleProcessing","Region","Sex"),
     do.scale = T, 
     anno_columns = c("Experiment","Sex","Condition","TimePoint", "SampleProcessing","Region"), 
     cluster_rows = T, main = paste("sig genes in",i ,sep ="_"),
     assayName = "corrected"))
print(topTags(Results[[i]],p.value = 0.05, n = 30))
  }else{
  print(paste(i, "had no significant entries", sep = " "))
  print(topTags(Results[[i]], n = 30))
}
}

tags <- topTags(Results[["FullModel"]],p.value = 0.05, n = 20)$table

radial.df <- data.frame(name = rownames(tags), logFC = tags$logFC.CB.ConditionSwim, Experiment = "CMV TRAP")
radial.df <- rbind(radial.df, data.frame(name = rownames(tags), logFC = tags$logFC.CB.ConditionSwim + tags$logFC.CB.ConditionSwim.CB.SampleProcessing2CAMK2A_TRAP, Experiment = "Camk2a TRAP"))
radial.df <- rbind(radial.df, data.frame(name = rownames(tags), logFC = tags$logFC.CB.ConditionSwim + tags$logFC.CB.ConditionSwim.CB.SampleProcessing2VIAAT_TRAP, Experiment = "Viaat TRAP"))
radial.df$direction = ifelse(radial.df$logFC >= 0, "up-regulated", "down-regulated")
radial.df$logFC <- abs(radial.df$logFC)
radial.df$Experiment <- as.factor(radial.df$Experiment)
radial.df$Experiment <- relevel(radial.df$Experiment, "CMV TRAP")

ggplot(radial.df, aes(x=name, y=logFC, fill=name, color = direction)) +
    geom_bar(stat='identity') + facet_wrap(~Experiment) + coord_polar() + scale_color_manual(values= c("red","black")) + theme_bw() + ggtitle("top 20 genes")
```
