---
title: "stress snRNAseq - Exploration"
author: "Pierre-Luc Germain"
date: "9/3/2020"
output:
  html_document:
    highlight: tango
    code_folding:	hide
---

<style>
  #content {
    max-width: 1200px !important;
  }
</style>

```{r}
suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(scater)
  library(cowplot)
  library(muscat)
  library(SEtools)
  library(ComplexHeatmap)
})
```

```{r}
sce <- readRDS("SCE.clust.rds")
```


## Library sizes

```{r, fig.height=4, fig.width=10}
sce$log_counts <- log(sce$sum)
plot_grid(
  plotReducedDim(sce, dimred = "TSNE", colour_by="log_counts"),
  plotReducedDim(sce, dimred = "UMAP", colour_by="log_counts"),
  nrow=1)
```

## Putative doublets

```{r, fig.height=8, fig.width=10}
plot_grid(
  plotReducedDim(sce, dimred = "TSNE", colour_by="scDblFinder.ratio"),
  plotReducedDim(sce, dimred = "TSNE", colour_by="scDblFinder.class"),
  plotReducedDim(sce, dimred = "UMAP", colour_by="scDblFinder.ratio"),
  plotReducedDim(sce, dimred = "UMAP", colour_by="scDblFinder.class"),
  nrow=2)
```

Some doublets are very clearly identifiable; in addition, there is a whole set of cells with (left on the UMAP) with suspiciously high ratio, we should keep an eye on those.

```{r}
sce <- sce[,sce$scDblFinder.class=="singlet"]
```


<br/><br/>

*** 

# Clustering

## Identifiability

```{r}
cs <- metadata(sce)$cluster.stability
resolutions <- sort(as.numeric(names(cs$global)))
boxplot(cs$perCluster[as.character(resolutions)], xlab="Leiden resolution", ylab="Misclassification rate")
lines(cs$global[as.character(resolutions)], lwd=2, col="blue")
```


## Clustering resolution {.tabset}

```{r}
genes <- list(
  astrocytes = c("Aqp4", "Gfap", "Fgfr3","Dio2"),
  endothelial = c("Cldn5","Nostrin","Flt1"),
  microglia = c("C1qb","Tyrobp","Spp1","P2ry12", "Csf1r", "Irf8"),
  neuronal = c("Snap25", "Stmn2", "Syn1", "Rbfox3", "Dlg4"),
  neuronal_excitatory = c("Slc17a7","Camk2a","Grin2b","Fezf2"),
  neuronal_inhibitory = c("Gad1","Lhx6","Adarb2"),
  oligodendrocyte = c("Opalin","Plp1","Mag","Mog"),
  OPC = c("Pdgfra","Sox6","Bcan"),
  CPE = "Ttr",
  misc=c("Eps8","Lama2")
)
genes2 <- lapply(genes, FUN=function(x) row.names(sce)[rowData(sce)$Symbol %in% x])
mcols <- pipeComp::getQualitativePalette(length(genes2))
names(mcols) <- names(genes2)

prepPB <- function(sce, cl, genes2){
  pb <- flattenPB(aggregateData(sce, "counts", by=c(cl,"sample_id")))
  pb$logReadCount <- log(colSums(assay(pb)))
  rowData(pb)$marker <- factor(NA, levels=names(genes2))
  for(f in setdiff(names(genes2),"misc")) rowData(pb)[genes2[[f]],"marker"] <- f
  metadata(pb)$anno_colors <- list(marker=mcols)
  pb
}
```

```{r results="asis", echo=FALSE, fig.height=4, fig.width=8}
if(is.null(resolutions)) resolutions <- sort(as.numeric(gsub("SCT_snn_res\\.","",
                 grep("SCT_snn_res", colnames(colData(sce)), value=TRUE))))
for(i in resolutions){
  cl <- paste0("SCT_snn_res.",i)
  sce[[cl]] <- droplevels(sce[[cl]])
  cat("### Resolution=",i,"\n\n")
  print(plot_grid(
      plotReducedDim(sce, dimred = "TSNE", colour_by=cl, text_by=cl) + theme(legend.position="none"),
      plotReducedDim(sce, dimred = "UMAP", colour_by=cl, text_by=cl) + theme(legend.position="none"),
      nrow=1, labels=rep(as.character(i),2)))
  
  cat("\n\n")
}
```

## Clustering resolution - markers (pseudobulk) {.tabset}


```{r results="asis", echo=FALSE, fig.height=8, fig.width=8}
for(i in sort(resolutions)){
  cl <- paste0("SCT_snn_res.",i)
  cat("### Resolution=",i,"\n\n")
  pb <- prepPB(sce, cl, genes2)
  draw(sechm(pb, unlist(genes2), assayName="logcpm", do.scale=TRUE, anno_columns=c("n_cells","logReadCount"), gaps_at = c("cluster_id"), anno_rows = "marker", gaps_row=rowData(pb)$marker, row_names_gp = gpar(fontsize = 10)), annotation_legend_side="bottom", heatmap_legend_side="bottom", merge_legends=TRUE)
  cat("\n\n")
}
```

<br/><br/>
***

# Expression of known markers

```{r, fig.height=11, fig.width=10}
plot_grid(
  plotReducedDim(sce, dimred = "TSNE", colour_by="SCT_snn_res.0.5", text_by = "SCT_snn_res.0.5") + theme(legend.position="none") + ggtitle("Clustering at res=0.5"),
  plotReducedDim(sce, dimred = "TSNE", colour_by="ENSMUSG00000054252.Fgfr3"),
  plotReducedDim(sce, dimred = "TSNE", colour_by="ENSMUSG00000027273.Snap25"),
  plotReducedDim(sce, dimred = "TSNE", colour_by="ENSMUSG00000030209.Grin2b"),
  plotReducedDim(sce, dimred = "TSNE", colour_by="ENSMUSG00000050121.Opalin"),
  plotReducedDim(sce, dimred = "TSNE", colour_by="ENSMUSG00000041378.Cldn5"),
  nrow=3)
```
